---
title: "Wood spectrum pre-processing for p279 population in 2020"
author: "V Segura, C Brault, M Ecarnot, L Le Cunff"
date: "`r format(Sys.time(), '%d/%m/%Y %H:%M:%S')`"
output:
  html_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 3
  word_document:
    toc: yes
    toc_depth: '3'
lang: en
colorlinks: yes
urlcolor: blue
---

<!--
This R chunk is used to set up some options.
-->
```{r setup, include=FALSE}
##`
## https://github.com/vspinu/polymode/issues/147#issuecomment-399745611
R.v.maj <- as.numeric(R.version$major)
R.v.min.1 <- as.numeric(strsplit(R.version$minor, "\\.")[[1]][1])
if(R.v.maj < 2 || (R.v.maj == 2 && R.v.min.1 < 15))
  stop("requires R >= 2.15", call.=FALSE)

suppressPackageStartupMessages(library(knitr))
opts_chunk$set(echo=TRUE, warning=TRUE, message=TRUE, cache=FALSE, fig.align="center")
opts_knit$set(progress=TRUE, verbose=TRUE)
```


# Overview

This project is about **phenomic prediction** i.e using Near Infra-Red Spectrum of biological tissues of different varieties to predict phenotypes instead of genotyping (Rincent et al, 2018).
We took spectra on 3 grapevine populations:

* bi-parental cross Syrah x Grenache, with 190 genotypes (SG)

* a core collection of 279 genotypes (CC_279)

* A p279 facility with 5 parents and about 700 genotypes (diallel) 

In this script, we pre-process raw spectrum data.

```{r overview}
project.name <- "phenomic_pred"
# if(Sys.info()["user"] %in% c("brault", "charlotte")){
#   project.dir <- paste0("~/work/", project.name)
# } 
project.dir <- "~/breeder_equation_project/data/interest_of_phenomic_prediction_as_an_alternative_to_genomic_predition/script/phenomic_pred/"
data.dir <- paste0(project.dir, "/data")
results.dir <- paste0(project.dir, "/results")
src.dir <- paste0(project.dir,"/src")
```

This document will also require external packages to be available:


```{r load_pkg}
suppressPackageStartupMessages(library(readxl))
suppressPackageStartupMessages(library(data.table))
suppressPackageStartupMessages(library(apercu))
suppressPackageStartupMessages(library(RColorBrewer))
suppressPackageStartupMessages(library(signal))
suppressPackageStartupMessages(library(prospectr))
suppressPackageStartupMessages(library(ade4))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(viridis))
suppressPackageStartupMessages(library(VCA))
suppressPackageStartupMessages(library(parallel))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(rnirs)) # on GitHub
suppressPackageStartupMessages(library(doParallel)) # for parallelization
suppressPackageStartupMessages(library(foreach)) # for parallelization
suppressPackageStartupMessages(library(snow)) # for creating cluster
source(file = paste0(src.dir,"/PS_functions.R")) # for reading asd files
source(paste0(src.dir,"/ortho.nirs.R")) # ortho.nirs

nb.cores <- 5
cl <- makeSOCKcluster(nb.cores)
registerDoParallel(cl)
```

This R chunk is used to assess how much time it takes to execute the R code in this document until the end:
```{r time_0}
t0 <- proc.time()
```


# Load data

## NIRS spectrum

Wood spectra for p279 population and bloc A and B together.
```{r read ASD, warning=FALSE}
nirs <- asd_read_dir(directory=paste0(data.dir, "/spectra/p279_wood_2020"))
dim(nirs) # 1896 x 2151
ap(nirs)
length(unique(rownames(nirs))) #1896
```


## Link to genotype information

```{r load geno info}
p2f <- paste0(data.dir, "/spectra/tableau manip NIRS-bois_final.xlsx")
tools::md5sum(p2f)
dat_geno <- read_xlsx(p2f,sheet = "CC279", na = "x")
dat_geno <- as.data.frame(dat_geno)
dim(dat_geno) # 510 x 12
head(dat_geno)
str(dat_geno)

dat_geno2 <- reshape2::melt(dat_geno,id.vars=c("Pop", "Bloc", "Rang", "Plac",
                                               "Id", "Pop-id", "Bl-Rg-Pl","Remarque"),
                            variable.name="NIRS.origin", value.name="NIRS.nb")
head(dat_geno2)
dat_geno2$Id <- as.factor(dat_geno2$Id)
dat_geno2$Bloc <- as.factor(dat_geno2$Bloc)
dat_geno2$NIRS.nb <- as.numeric(dat_geno2$NIRS.nb)
# remove missing values
dat_geno2 <- dat_geno2[!is.na(dat_geno2$NIRS.nb),]
```


## Link NIRS to sample information

```{r load sample info}
sample.info <- data.frame("NIRS.id" = rownames(nirs),stringsAsFactors = FALSE)
sample.info$NIRS.nb <- as.numeric(substring(sample.info$NIRS.id, 9))
head(sample.info)

summary(sample.info$NIRS.nb)  # min 1 max 1896
summary(dat_geno2$NIRS.nb) # min 1 max 1896

stopifnot(length(setdiff(sample.info$NIRS.nb, dat_geno2$NIRS.nb)) == 0,
          length(setdiff(dat_geno2$NIRS.nb,sample.info$NIRS.nb)) == 0)

length(unique(sample.info$NIRS.nb)) # 1896
length(unique(dat_geno2$NIRS.nb)) # 1896
stopifnot(!any(duplicated(sample.info$NIRS.nb)),
          !any(duplicated(dat_geno2$NIRS.nb)),
          nrow(dat_geno2) == length(unique(dat_geno2$NIRS.nb)))

sample.info <- merge(sample.info, dat_geno2, by="NIRS.nb", all=FALSE) 
dim(sample.info) # 1896 x 11
identical(rownames(nirs), sample.info$NIRS.id)
sample.info$Id_bloc <- paste0(sample.info$Id, "_", sample.info$Bloc)
```


## Identify controls and p279 subpopulations

```{r controls identification}
subpop <- as.data.frame(read_xlsx(paste0(data.dir, "/genotypic/CC279_GBS/origin_CC279.xlsx")))
sample.info <- merge(sample.info, subpop, by.x="Id", by.y="code.intro", all=FALSE)

infos_geno <- as.data.frame(read_xlsx(paste0(data.dir, "/genotypic/CC279_GBS/ids_279.xlsx"), na = ""))
sample.info <- merge(sample.info, infos_geno, by.x="cultivar.number", by.y="CodeVar_TL")

sample.infoCtl <- sample.info[sample.info$Id %in% c("TEMOIN"),]
sample.infop279 <- sample.info[!sample.info$Id %in% c("TEMOIN"),]

dim(sample.infoCtl)
dim(sample.infop279)

head(sample.infop279)
```

# Read spectrum

## Plot raw spectrum

```{r graph nirs, fig.height=4.5, fig.width = 7}
matplot(x = as.integer(colnames(nirs)),
        y = t(nirs),
        type = 'l', lty = 1,
        xlab = "Wavelength (nm)", ylab = "Reflectance",
        main = "Wood NIRS p279 2020 - raw spectrum")
```

## Remove jump corresponding to change in detector

We use the following function *adj_asd*:

```{r spliceCorrection, fig.height=4.5, fig.width = 7}
nirs <- adj_asd(Xi = nirs, iadj  = c(651, 1451))
matplot(x = as.integer(colnames(nirs)),
        y = t(nirs),
        type = 'l', lty = 1,
        xlab = "Wavelength (nm)", ylab = "Reflectance",
        main = "Wood NIRS p279 2020 - Remove jump")
```

## Cut the end of the sprectrum (50 nm visible side)

```{r spectra cut, fig.height=4.5, fig.width = 7}
nirs.cut <- nirs[, as.integer(colnames(nirs)) >= 400]
dim(nirs.cut)
matplot(x = as.integer(colnames(nirs.cut)),
        y = t(nirs.cut),
        type = 'l', lty = 1,
        xlab = "wavelength (nm)", ylab = "Reflectance",
        main = "Wood NIRS p279 2020 - cut")
```

## Average of 4 scans done on each sample

For each genotype in one block, there are 4 scans : two sides of each shoot and two shoots.

```{r nirs.ok, fig.height=4.5, fig.width = 7}
# for p279, one genotype per bloc
nirs.meanp279 <- apply(nirs.cut[sample.infop279$NIRS.nb,], 2, function(x) {
  tapply(X = x, INDEX = sample.infop279$Id_bloc, FUN = mean)
})
# For controls, several controls per bloc
nirs.meanCtl <- apply(nirs.cut[sample.infoCtl$NIRS.nb,], 2, function(x) {
  tapply(X = x, INDEX = sample.infoCtl$Id_bloc, FUN = mean)
})
nirs.mean <- rbind(nirs.meanp279, nirs.meanCtl)


matplot(x = as.integer(colnames(nirs.mean)),
        y = t(nirs.mean),
        type = 'l', lty = 1,
        xlab = "Wavelength (nm)", ylab = "Reflectance",
        main = "Wood NIRS p279 2020 - mean")
```

# Correct for work session effect with REP-ASCA

## Estimate correction on the whole population

Use `ortho.nirs` function which applies REP-ASCA correction to maximize the genotypic effect.
```{r apply orthogonalization}
dim(nirs.mean)

# create a new data frame with one line per Id_bloc
idx <- !duplicated(sample.info$Id_bloc)
sample.info.mean <- data.frame(Id_bloc=sample.info$Id_bloc[idx],
                               Id=sample.info$Id[idx], NIRS=I(nirs.mean),
                               stringsAsFactors=FALSE)
dim(sample.info.mean)

out <- ortho.nirs(data=sample.info.mean, SpName="NIRS", corr.effect=NULL,
                  max.effect="Id", ncomp=1, plot=TRUE, verbose=1)
head(out$loadings)
dim(out$loadings)
dim(out$corrected)
nirs.mean.corr <- out$corrected

matplot(x = as.integer(colnames(nirs.mean.corr)),
        y = t(nirs.mean.corr),
        type = 'l', lty = 1,
        xlab = "Wavelength (nm)", ylab = "Reflectance",
        main = "Wood NIRS p279 2020 - mean after correction")
nirs.mean.corr <- nirs.mean # don't apply correction for the moment
```

<!-- ### Minimize inter-session variance -->

<!-- ```{r} -->
<!-- out.session <- ortho.nirs(data=sample.info.mean, SpName="NIRS", corr.effect="session", -->
<!--                           max.effect=NULL, ncomp=1, plot=TRUE, verbose=1) -->
<!-- # type.corr <- "n1-session" -->
<!-- # head(out.session$loadings) -->
<!-- # dim(out.session$loadings) -->
<!-- # dim(out.session$corrected) -->
<!-- # nirs.mean.corr <- out.session$corrected -->
<!-- #  -->
<!-- # matplot(x = as.integer(colnames(out.session$corrected)), -->
<!-- #         y = t(out.session$corrected), -->
<!-- #         type = 'l', lty = 1, -->
<!-- #         xlab = "Wavelength (nm)", ylab = "Reflectance", -->
<!-- #         main = "Wood NIRS p279 - mean after correction") -->

<!-- ``` -->

# Pre-process spectrum

## Smoothing

Smoothing of spectra because some area are noisy and it could pose a challenge for other pre-process (derivative).
Apply this on corrected spectra.

```{r smooth, fig.height=5, fig.width=7}
nirs.smooth <- t(apply(nirs.mean.corr, 1, function(x) {
  sgolayfilt(x, p = 1, n = 21, m = 0, ts = 1)}))
colnames(nirs.smooth) <- colnames(nirs.mean.corr)
matplot(x = as.integer(colnames(nirs.smooth)),
        y = t(nirs.smooth),
        type = 'l', lty = 1,
        xlab = "Wavelength (nm)", ylab = "Reflectance",
        main = "Wood NIRS p279 2020 - smooth")
```

## Normalisation (SNV)

```{r norm, fig.height=4, fig.width=7}
nirs.snv <- t(scale(t(nirs.smooth)))

matplot(x = as.integer(colnames(nirs.snv)),
        y = t(nirs.snv),
        type = 'l', lty = 1,
        xlab = "Wavelength (nm)", ylab = "Reflectance",
        main = "Wood NIRS p279 2020 - snv")
```

## Detrend

```{r dt, fig.height=4, fig.width=7}
nirs.dt <- prospectr::detrend(X = nirs.smooth,
                              wav = as.integer(colnames(nirs.smooth)))

matplot(x = as.integer(colnames(nirs.dt)),
        y = t(nirs.dt),
        type = 'l', lty = 1,
        xlab = "Wavelength (nm)", ylab = "Reflectance",
        main = "Wood NIRS p279 2020 - detrend")
```

## Derivative 1

Derivative done on SNV spectrum.

```{r der1, fig.height=4, fig.width=7}
nirs.der1 <- t(apply(nirs.snv, 1, function(x) {
  sgolayfilt(x, p = 2, n = 41, m = 1, ts = 1)}))

colnames(nirs.der1) <- colnames(nirs.snv)
matplot(x = as.integer(colnames(nirs.der1)),
        y = t(nirs.der1),
        type = 'l', lty = 1,
        xlab = "Wavelength (nm)", ylab = "Reflectance",
        main = "Wood NIRS p279 2020 - der1")
```

## Derivative 2

```{r der2, fig.height=4, fig.width=7}
nirs.der2 <- t(apply(nirs.snv, 1, function(x) {
  sgolayfilt(x, p = 3, n = 81, m = 2, ts = 1)}))

colnames(nirs.der2) <- colnames(nirs.snv)
matplot(x = as.integer(colnames(nirs.der2)),
        y = t(nirs.der2),
        type = 'l', lty = 1,
        xlab = "Wavelength (nm)", ylab = "Reflectance",
        main = "Wood NIRS p279 2020 - der2")
```

## List of pre-process spectra

For the p279

```{r spectra list p279}
nirs.list <- list("raw" = nirs.mean, "smooth" = nirs.smooth,
                  "snv" = nirs.snv, "dt" = nirs.dt,
                  "der1" = nirs.der1, "der2" = nirs.der2)
nirs.listp279 <- lapply(nirs.list, function(x) {
  merge(sample.infop279, data.frame("id" = rownames(x), "NIRS" = I(x)),
        by.x="Id_bloc", by.y = "id", all=FALSE)
})
nirs.listp279 <- lapply(nirs.listp279, function(x){
  x[!duplicated(x$Id_bloc),]})
lapply(nirs.listp279, dim)




# 
# ## On corrected NIRS
# 
# nirs.list_corr <- list("snv corr" = nirs.snv_corr, "dt corr" = nirs.dt_corr,
#                   "der1 corr" = nirs.der1_corr, "der2 corr" = nirs.der2_corr)
# 
# nirs.listp279_corr <- lapply(nirs.list_corr, function(x) {
#   merge(sample.infop279, data.frame("id" = rownames(x), "NIRS" = I(x)),
#         by.x="Id_bloc", by.y = "id", all=FALSE)
# })
# nirs.listp279_corr <- lapply(nirs.listp279_corr, function(x){
#   x[!duplicated(x$Id_bloc),]})
# lapply(nirs.listp279_corr, dim)
```

For controls

```{r spectra list ctl}
nirs.listCtl <- lapply(nirs.list, function(x) {
  merge(sample.infoCtl, data.frame("id" = rownames(x), "NIRS" = I(x)),
        by.x="Id_bloc", by.y = "id", all=FALSE)
})
nirs.listCtl <- lapply(nirs.listCtl, function(x){
  x[!duplicated(x$Id_bloc),]})
lapply(nirs.listCtl, dim)

# #On corrected NIRS
# 
# nirs.listCtl_corr <- lapply(nirs.list_corr, function(x) {
#   merge(sample.infoCtl, data.frame("id" = rownames(x), "NIRS" = I(x)),
#         by.x="Id_bloc", by.y = "id", all=FALSE)
# })
# nirs.listCtl_corr <- lapply(nirs.listCtl_corr, function(x){
#   x[!duplicated(x$Id_bloc),]})
# lapply(nirs.listCtl_corr, dim)
```

For all

```{r spectra list all}
nirs.list <- lapply(nirs.list, function(x) {
  merge(sample.info, data.frame("id" = rownames(x), "NIRS" = I(x)),
        by.x="Id_bloc", by.y = "id", all=FALSE)
})
nirs.list <- lapply(nirs.list, function(x){
  x[!duplicated(x$Id_bloc),]})
lapply(nirs.list, dim)

# save spectra
p2f <- paste0(results.dir,"/spectrum_correction/wood_p279_2020_nirs-list-2-blocks.Rdata")
save(nirs.list, file=p2f)

# #On corrected NIRS
# 
# nirs.list_corr <- lapply(nirs.list_corr, function(x) {
#   merge(sample.info, data.frame("id" = rownames(x), "NIRS" = I(x)),
#         by.x="Id_bloc", by.y = "id", all=FALSE)
# })
# nirs.list_corr <- lapply(nirs.list_corr, function(x){
#   x[!duplicated(x$Id_bloc),]})
# lapply(nirs.list_corr, dim)
```

```{r echo = F}
rm(nirs.cut, nirs.smooth, nirs.snv, nirs.dt, nirs.der1, nirs.der2)
```

# Exploratory analysis (PCA)

## PCA on all population

```{r pca ade4 all}
data.ade4 <- lapply(nirs.list, function(x){
  list("info" = data.frame(x[, - which(colnames(x) == "NIRS")]),
       "NIRS" = as.data.frame(matrix(x$NIRS, nrow = nrow(x$NIRS),
                                     ncol = ncol(x$NIRS))))})
pca.ade4 <- lapply(data.ade4, function(x){
  dudi.pca(x$NIRS, center=TRUE, scale=TRUE, scannf=F, nf=20)})

# # On corrected NIRS
# 
# data.ade4_corr <- lapply(nirs.list_corr, function(x){
#   list("info" = data.frame(x[, - which(colnames(x) == "NIRS")]),
#        "NIRS" = as.data.frame(matrix(x$NIRS, nrow = nrow(x$NIRS),
#                                      ncol = ncol(x$NIRS))))})
# pca.ade4_corr <- lapply(data.ade4_corr, function(x){
#   dudi.pca(x$NIRS, center=TRUE, scale=TRUE, scannf=F, nf=20)})
```


### Inertia

```{r percvar pca all, fig.height=10, fig.width=6}
par(mfrow = c(3, 2))

pve.pca <- sapply(sapply(pca.ade4, function(x){x$eig / sum(x$eig) * 100}),
                  function(y){y[1:10]})
rownames(pve.pca) <- paste("PC", 1:nrow(pve.pca), sep = "")
for (i in 1:ncol(pve.pca)){
  barplot(pve.pca[, i], col = "grey",
          main = colnames(pve.pca)[i],
          ylab="% variance", ylim = c(0, round(max(pve.pca))))
}

# #On corrected NIRS
# 
# pve.pca_corr <- sapply(sapply(pca.ade4_corr, function(x){x$eig / sum(x$eig) * 100}),
#                   function(y){y[1:10]})
# rownames(pve.pca_corr) <- paste("PC", 1:nrow(pve.pca_corr), sep = "")
# for (i in 1:ncol(pve.pca_corr)){
#   barplot(pve.pca_corr[, i], col = "grey",
#           main = colnames(pve.pca_corr)[i],
#           ylab="% variance", ylim = c(0, round(max(pve.pca_corr))))
# }
```

### Projection of spectrum on the first 2 axis

```{r slabel pca all, fig.height=10, fig.width=6}
par(mfrow = c(3, 2))
for (i in 1:length(pca.ade4)){
  s.label(pca.ade4[[i]]$li, label = data.ade4[[i]]$info$Id_bloc,
          sub = names(pca.ade4)[i], clabel = 0.6)
}

# #On corrected NIRS
# 
# for (i in 1:length(pca.ade4_corr)){
#   s.label(pca.ade4_corr[[i]]$li, label = data.ade4_corr[[i]]$info$Id_bloc,
#           sub = names(pca.ade4_corr)[i], clabel = 0.6)
# }
```



### With color per bloc

```{r sclass pca all bloc, fig.height=8, fig.width=6}
# col.bloc <- grepl("BlocA", rownames(nirs.mean))
# col.bloc <- ifelse(col.bloc, "dodgerblue", "indianred")
par(mfrow = c(3, 2))

for (i in (1:length(pca.ade4))){
  data.ade4[[i]]$info$Bloc <- as.factor(data.ade4[[i]]$info$Bloc)
  s.class(dfxy = pca.ade4[[i]]$li, fac = data.ade4[[i]]$info$Bloc,
          xax = 1, yax = 2, col = c("dodgerblue", "indianred"),
          sub = names(pca.ade4)[i], csub=2, possub="topleft")
}

# 
# #On corrected NIRS
# 
# for (i in (1:length(pca.ade4_corr))){
#   data.ade4_corr[[i]]$info$Bloc <- as.factor(data.ade4_corr[[i]]$info$Bloc)
#   s.class(dfxy = pca.ade4_corr[[i]]$li, fac = data.ade4_corr[[i]]$info$Bloc,
#           xax = 1, yax = 2, col = col,
#           sub = names(pca.ade4_corr)[i], csub=2, possub="topleft")
# }
```

### With color per subpopulation
TODO
```{r sclass pca all cross, fig.height=8, fig.width=6}
# # col <- c("dodgerblue", "indianred", "seagreen", "gold3")
# par(mfrow = c(3, 2))
# for (i in (1:length(pca.ade4))){
#   data.ade4[[i]]$info$cross <- as.factor(data.ade4[[i]]$info$cross)
#   s.class(dfxy = pca.ade4[[i]]$li, fac = data.ade4[[i]]$info$cross,
#           xax = 1, yax = 2, col = colors_cross(data.ade4[[i]]$info$cross),
#           sub = names(pca.ade4)[i], csub=2, possub="topleft")
# }

# #On corrected NIRS
# 
# for (i in (1:length(pca.ade4_corr))){
#   data.ade4_corr[[i]]$info$status <- as.factor(data.ade4_corr[[i]]$info$status)
#   s.class(dfxy = pca.ade4_corr[[i]]$li, fac = data.ade4_corr[[i]]$info$status,
#           xax = 1, yax = 2, col = col,
#           sub = names(pca.ade4_corr)[i], csub=2, possub="topleft")
# }
```


<!-- ### With color if uncertainty/remarks -->

<!-- ```{r sclass pca all remarks, fig.height=20, fig.width=6} -->
<!-- unique(data.ade4$raw$info$Remarque) -->
<!-- col <- brewer.pal(length(unique(data.ade4$raw$info$Remarque)), -->
<!--                       name = "Set1") -->
<!-- par(mfrow = c(6, 2)) -->
<!-- for (i in (1:length(pca.ade4))){ -->
<!--   data.ade4[[i]]$info$Remarque[is.na(data.ade4[[i]]$info$Remarque)] <- "OK" -->
<!--   data.ade4[[i]]$info$Remarque <- as.factor(data.ade4[[i]]$info$Remarque) -->
<!--   s.class(dfxy = pca.ade4[[i]]$li, fac = data.ade4[[i]]$info$Remarque, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- #On corrected NIRS -->

<!-- for (i in (1:length(pca.ade4_corr))){ -->
<!--   data.ade4_corr[[i]]$info$Remarque[is.na(data.ade4_corr[[i]]$info$Remarque)] <- "OK" -->
<!--   data.ade4_corr[[i]]$info$Remarque <- as.factor(data.ade4_corr[[i]]$info$Remarque) -->
<!--   s.class(dfxy = pca.ade4_corr[[i]]$li, fac = data.ade4_corr[[i]]$info$Remarque, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4_corr)[i], csub=2, possub="topleft") -->
<!-- } -->
<!-- ``` -->





### With color per rank (ranks are chronologically colored)

```{r sclass pca p279 rank, fig.height=8, fig.width=6}
sample.infotri <- arrange(sample.info, sample.info$NIRS.nb)
sample.info$NIRS.nb <- as.factor(sample.info$NIRS.nb)
rank <- unique(sample.infotri$Rang)
col <- viridis::viridis(length(rank))
par(mfrow = c(3, 2))


for (i in (1:length(pca.ade4))){
  data.ade4[[i]]$info$Rang <- factor(data.ade4[[i]]$info$Rang,
                                        levels=rank)#[order(rank)])
  s.class(dfxy = pca.ade4[[i]]$li, fac = data.ade4[[i]]$info$Rang,
          xax = 1, yax = 2, col = col,
          sub = names(pca.ade4)[i], csub=2, possub="topleft")
}
# 
# #On corrected NIRS
# 
# for (i in (1:length(pca.ade4_corr))){
#   data.ade4_corr[[i]]$info$Rang <- factor(data.ade4_corr[[i]]$info$Rang,
#                                         levels=rank)#[order(rank)])
#   s.class(dfxy = pca.ade4_corr[[i]]$li, fac = data.ade4_corr[[i]]$info$Rang,
#           xax = 1, yax = 2, col = col,
#           sub = names(pca.ade4_corr)[i], csub=2, possub="topleft")
# }
```

<!-- ### With color per work sessions -->

<!-- ```{r sclass pca p279 work sessions, fig.height=10, fig.width=6} -->
<!-- col <- c("hotpink2", "aquamarine2") -->

<!-- par(mfrow = c(3, 2)) -->

<!-- for (i in (1:length(pca.ade4))){ -->
<!--   data.ade4[[i]]$info$Manip_Session <- as.factor(data.ade4[[i]]$info$Manip_Session) -->
<!--   s.class(dfxy = pca.ade4[[i]]$li, fac = data.ade4[[i]]$info$Manip_Session, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- # #On corrected NIRS -->
<!-- #  -->
<!-- # for (i in (1:length(pca.ade4_corr))){ -->
<!-- #   data.ade4_corr[[i]]$info$Manip_Session <- as.factor(data.ade4_corr[[i]]$info$Manip_Session) -->
<!-- #   s.class(dfxy = pca.ade4_corr[[i]]$li, fac = data.ade4_corr[[i]]$info$Manip_Session, -->
<!-- #           xax = 1, yax = 2, col = col, -->
<!-- #           sub = names(pca.ade4_corr)[i], csub=2, possub="topleft") -->
<!-- # } -->
<!-- ``` -->


<!-- ## PCA on p279 -->

<!-- ```{r pca ade4 p279} -->
<!-- data.ade4p279 <- lapply(nirs.listp279, function(x){ -->
<!--   list("info" = data.frame(x[, - which(colnames(x) == "NIRS")]), -->
<!--        "NIRS" = as.data.frame(matrix(x$NIRS, nrow = nrow(x$NIRS), -->
<!--                                      ncol = ncol(x$NIRS))))}) -->
<!-- pca.ade4p279 <- lapply(data.ade4p279, function(x){ -->
<!--   dudi.pca(x$NIRS, center=TRUE, scale=TRUE, scannf=F, nf=20)}) -->

<!-- # #On corrected NIRS -->
<!-- #  -->
<!-- # data.ade4p279_corr <- lapply(nirs.listp279_corr, function(x){ -->
<!-- #   list("info" = data.frame(x[, - which(colnames(x) == "NIRS")]), -->
<!-- #        "NIRS" = as.data.frame(matrix(x$NIRS, nrow = nrow(x$NIRS), -->
<!-- #                                      ncol = ncol(x$NIRS))))}) -->
<!-- # pca.ade4p279_corr <- lapply(data.ade4p279_corr, function(x){ -->
<!-- #   dudi.pca(x$NIRS, center=TRUE, scale=TRUE, scannf=F, nf=20)}) -->
<!-- ``` -->


<!-- ### Inertia -->

<!-- ```{r percvar pca p279, fig.height=20, fig.width=6} -->
<!-- par(mfrow = c(6, 2)) -->

<!-- pve.pcap279 <- sapply(sapply(pca.ade4p279, function(x){x$eig / sum(x$eig) * 100}), -->
<!--                   function(y){y[1:10]}) -->
<!-- rownames(pve.pcap279) <- paste("PC", 1:nrow(pve.pcap279), sep = "") -->

<!-- for (i in 1:ncol(pve.pcap279)){ -->
<!--   barplot(pve.pcap279[, i], col = "grey", -->
<!--           main = colnames(pve.pcap279)[i], -->
<!--           ylab="% variance", ylim = c(0, round(max(pve.pcap279)))) -->
<!-- } -->

<!-- #On corrected NIRS -->

<!-- pve.pcap279_corr <- sapply(sapply(pca.ade4p279_corr, function(x){x$eig / sum(x$eig) * 100}), -->
<!--                   function(y){y[1:10]}) -->
<!-- rownames(pve.pcap279_corr) <- paste("PC", 1:nrow(pve.pcap279_corr), sep = "") -->
<!-- for (i in 1:ncol(pve.pcap279_corr)){ -->
<!--   barplot(pve.pcap279_corr[, i], col = "grey", -->
<!--           main = colnames(pve.pcap279_corr)[i], -->
<!--           ylab="% variance", ylim = c(0, round(max(pve.pcap279_corr)))) -->
<!-- } -->
<!-- ``` -->

<!-- ### Projection of spectrum on the first 2 axis -->

<!-- ```{r slabel pca p279, fig.height=20, fig.width=6} -->
<!-- par(mfrow = c(6, 2)) -->
<!-- for (i in 1:length(pca.ade4p279)){ -->
<!--   s.label(pca.ade4p279[[i]]$li, label = data.ade4p279[[i]]$info$Id_bloc, -->
<!--           sub = names(pca.ade4p279)[i], clabel = 0.6) -->
<!-- } -->

<!-- #On corrected NIRS -->

<!-- for (i in 1:length(pca.ade4p279_corr)){ -->
<!--   s.label(pca.ade4p279_corr[[i]]$li, label = data.ade4p279_corr[[i]]$info$Id_bloc, -->
<!--           sub = names(pca.ade4p279_corr)[i], clabel = 0.6) -->
<!-- } -->
<!-- ``` -->

<!-- ### With color per block -->

<!-- ```{r sclass pca p279 block, fig.height=20, fig.width=6} -->
<!-- col <- c("dodgerblue", "indianred") -->
<!-- par(mfrow = c(6, 2)) -->

<!-- for (i in (1:length(pca.ade4p279))){ -->
<!--   data.ade4p279[[i]]$info$Bloc <- as.factor(data.ade4p279[[i]]$info$Bloc) -->
<!--   s.class(dfxy = pca.ade4p279[[i]]$li, fac = data.ade4p279[[i]]$info$Bloc, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- #On corrected NIRS -->

<!-- for (i in (1:length(pca.ade4p279_corr))){ -->
<!--   data.ade4p279_corr[[i]]$info$Bloc <- as.factor(data.ade4p279_corr[[i]]$info$Bloc) -->
<!--   s.class(dfxy = pca.ade4p279_corr[[i]]$li, fac = data.ade4p279_corr[[i]]$info$Bloc, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279_corr)[i], csub=2, possub="topleft") -->
<!-- } -->
<!-- ``` -->
<!-- ### With color per berry color -->

<!-- ```{r sclass pca p279 berry color, fig.height=20, fig.width=6} -->
<!-- col <- col <- brewer.pal(length(unique(data.ade4p279$raw$info$Couleur)), -->
<!--                       name = "Dark2") -->
<!-- par(mfrow = c(6, 2)) -->

<!-- for (i in (1:length(pca.ade4p279))){ -->
<!--   data.ade4p279[[i]]$info$Couleur <- as.factor(data.ade4p279[[i]]$info$Couleur) -->
<!--   s.class(dfxy = pca.ade4p279[[i]]$li, fac = data.ade4p279[[i]]$info$Couleur, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- #On corrected NIRS -->

<!-- for (i in (1:length(pca.ade4p279_corr))){ -->
<!--   data.ade4p279_corr[[i]]$info$Couleur <- as.factor(data.ade4p279_corr[[i]]$info$Couleur) -->
<!--   s.class(dfxy = pca.ade4p279_corr[[i]]$li, fac = data.ade4p279_corr[[i]]$info$Couleur, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279_corr)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- ``` -->

<!-- ### With color per geographical origin -->

<!-- ```{r sclass pca p279 origin, fig.height=20, fig.width=6} -->
<!-- col <- c("hotpink3", "aquamarine2", "yellow2") -->
<!-- par(mfrow = c(6, 2)) -->

<!-- for (i in (1:length(pca.ade4p279))){ -->
<!--   data.ade4p279[[i]]$info$Origin <- as.factor(data.ade4p279[[i]]$info$Origin) -->
<!--   s.class(dfxy = pca.ade4p279[[i]]$li, fac = data.ade4p279[[i]]$info$Origin, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- #On corrected NIRS -->

<!-- for (i in (1:length(pca.ade4p279_corr))){ -->
<!--   data.ade4p279_corr[[i]]$info$Origin <- as.factor(data.ade4p279_corr[[i]]$info$Origin) -->
<!--   s.class(dfxy = pca.ade4p279_corr[[i]]$li, fac = data.ade4p279_corr[[i]]$info$Origin, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279_corr)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- ``` -->


<!-- ### With color per : Country, Group and Sub-group -->

<!-- ```{r sclass pca p279 country, fig.height=20, fig.width=6} -->
<!-- col <- pals::alphabet(length(unique(data.ade4p279$raw$info$Pays))) -->
<!-- unique(sample.info$Pays) -->
<!-- par(mfrow = c(6, 2)) -->
<!-- for (i in (1:length(pca.ade4p279))){ -->
<!--   data.ade4p279[[i]]$info$Pays <- as.factor(data.ade4p279[[i]]$info$Pays) -->
<!--   s.class(dfxy = pca.ade4p279[[i]]$li, fac = data.ade4p279[[i]]$info$Pays, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- #On corrected NIRS -->

<!-- for (i in (1:length(pca.ade4p279_corr))){ -->
<!--   data.ade4p279_corr[[i]]$info$Pays <- as.factor(data.ade4p279_corr[[i]]$info$Pays) -->
<!--   s.class(dfxy = pca.ade4p279_corr[[i]]$li, fac = data.ade4p279_corr[[i]]$info$Pays, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279_corr)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- ``` -->

<!-- ```{r sclass pca p279 group, fig.height=20, fig.width=6} -->
<!-- col <- brewer.pal(length(unique(data.ade4p279$raw$info$Grp_Geo)), -->
<!--                       name = "Set3") -->
<!-- unique(sample.info$Grp_Geo) -->
<!-- par(mfrow = c(6, 2)) -->

<!-- for (i in (1:length(pca.ade4p279))){ -->
<!--   data.ade4p279[[i]]$info$Grp_Geo <- as.factor(data.ade4p279[[i]]$info$Grp_Geo) -->
<!--   s.class(dfxy = pca.ade4p279[[i]]$li, fac = data.ade4p279[[i]]$info$Grp_Geo, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- #On corrected NIRS -->

<!-- for (i in (1:length(pca.ade4p279_corr))){ -->
<!--   data.ade4p279_corr[[i]]$info$Grp_Geo <- as.factor(data.ade4p279_corr[[i]]$info$Grp_Geo) -->
<!--   s.class(dfxy = pca.ade4p279_corr[[i]]$li, fac = data.ade4p279_corr[[i]]$info$Grp_Geo, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279_corr)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- ``` -->

<!-- ```{r sclass pca p279 sub-group, fig.height=20, fig.width=6} -->
<!-- col <- pals::alphabet(length(unique(data.ade4p279$raw$info$Sous_Grp_Geo))) -->
<!-- unique(sample.info$Sous_Grp_Geo) -->
<!-- par(mfrow = c(6, 2)) -->

<!-- for (i in (1:length(pca.ade4p279))){ -->
<!--   data.ade4p279[[i]]$info$Sous_Grp_Geo <- as.factor(data.ade4p279[[i]]$info$Sous_Grp_Geo) -->
<!--   s.class(dfxy = pca.ade4p279[[i]]$li, fac = data.ade4p279[[i]]$info$Sous_Grp_Geo, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- #On corrected NIRS -->

<!-- for (i in (1:length(pca.ade4p279_corr))){ -->
<!--   data.ade4p279_corr[[i]]$info$Sous_Grp_Geo <- as.factor(data.ade4p279_corr[[i]]$info$Sous_Grp_Geo) -->
<!--   s.class(dfxy = pca.ade4p279_corr[[i]]$li, fac = data.ade4p279_corr[[i]]$info$Sous_Grp_Geo, -->
<!--           xax = 1, yax = 2, col = col, -->
<!--           sub = names(pca.ade4p279_corr)[i], csub=2, possub="topleft") -->
<!-- } -->

<!-- ``` -->


# Variance partition

## Along the spectrum

### Apply ANOVA

Use a random effect model to partition the variance with `VCA` package.

Total variance is divided in variety effect (Id), bloc effect, and residual (corresponding to repetitions on same genotype)


```{r vca}
p2f <- paste0(results.dir, "/spectrum_correction/wood_p279_2020_outputVCA.Rdata")
if(!file.exists(p2f)){
system.time(output.vca <- mclapply(nirs.listp279, function(n) {
    output <- matrix(NA, nrow = 3, ncol = ncol(n$NIRS))
    colnames(output) <- colnames(n$NIRS)
    rownames(output) <- c("Id", "Id:Bloc", "error")
    for (i in 1:ncol(n$NIRS)) {
      tmp <- data.frame(n[, c("Id", "Bloc", "Id_bloc")],
                        y = as.vector(n$NIRS[, i]))

      vcmod <- anovaVCA(y ~ Id + Bloc, tmp, VarVC.method = "gb")
      output[, i] <- vcmod$aov.tab[2:4, "%Total"]
      rm(tmp, vcmod)
    }
    return(output)
  },
  mc.cores = 3))
  save(output.vca, file=p2f)
} else {
  load(p2f)
}
# 
# p2f_corr <- paste0(results.dir, "/outputVCA_p279_corr_all_n1-session.Rdata")
# if(!file.exists(p2f_corr)){
# system.time(output.vca_corr <- mclapply(nirs.listp279_corr, function(n) {
#     output_corr <- matrix(NA, nrow = 3, ncol = ncol(n$NIRS))
#     colnames(output_corr) <- colnames(n$NIRS)
#     rownames(output_corr) <- c("Id", "Id:Bloc", "error")
#     for (i in 1:ncol(n$NIRS)) {
#       tmp_corr <- data.frame(n[, c("Id", "Bloc", "Id_bloc")],
#                         y = as.vector(n$NIRS[, i]))
# 
#       vcmod_corr <- anovaVCA(y ~ Id + Bloc, tmp_corr, VarVC.method = "gb")
#       output_corr[, i] <- vcmod_corr$aov.tab[2:4, "%Total"]
#       rm(tmp_corr, vcmod_corr)
#     }
#     return(output_corr)
#   },
#   mc.cores = 1))
#   save(output.vca_corr, file=p2f_corr)
# } else {
#   load(p2f_corr)
# }
```

### Plot results

Plot variance partition along the spectrum

```{r plot vca, fig.height=15, fig.width=20}
par(mfrow = c(3, 2), mar = c(5, 5, 3, 4), xpd = TRUE)

varpropcum <- lapply(output.vca, function(x) {
  apply(x, 2, cumsum)
})
spectra_summary <- lapply(nirs.listp279, function(x) {
  data.frame("lambda" = as.integer(colnames(x$NIRS)),
             "mediane" = apply(x$NIRS, 2, median),
             "q1" = apply(x$NIRS, 2, function(x){quantile(x, 0.25)}),
             "q3" = apply(x$NIRS, 2, function(x){quantile(x, 0.75)}))
})

for (i in 1:length(spectra_summary)){
  plot(x = spectra_summary[[i]]$lambda, y = spectra_summary[[i]]$mediane, type = "n",
       ylim = c(0, 100), xlab = "wavelength (nm)",
       ylab = "Percentage of variances",
       main = names(varpropcum)[i])
  polygon(c(spectra_summary[[i]]$lambda, rev(spectra_summary[[i]]$lambda)),
          c(rep(0, ncol(varpropcum[[i]])), rev(varpropcum[[i]][1, ])),
          col = "brown1")
  polygon(c(spectra_summary[[i]]$lambda, rev(spectra_summary[[i]]$lambda)),
          c(varpropcum[[i]][1, ], rev(varpropcum[[i]][2, ])),
          col = "forestgreen")
  polygon(c(spectra_summary[[i]]$lambda, rev(spectra_summary[[i]]$lambda)),
          c(varpropcum[[i]][2, ], rev(varpropcum[[i]][3, ])),
          col = "dodgerblue4")
  par(new = TRUE)
  plot(spectra_summary[[i]]$lambda, spectra_summary[[i]]$mediane, type = "l",
       lty = 1, lwd = 1.5, ylim = range(nirs.listp279[[i]]$NIRS),
       axes = F, xlab = "", ylab = "", main = "", col = "black")
  polygon(c(spectra_summary[[i]]$lambda, rev(spectra_summary[[i]]$lambda)),
          c(spectra_summary[[i]]$q1, rev(spectra_summary[[i]]$q3)),
          col = "#C8C8C888", border = 0)
  axis(4)
  mtext("Refelctance", side = 4, line = 2.5, cex = 0.7)
}
legend(-100, 0.0019, c("res=bloc/genotype", "bloc", "genotype"),
       lty = 0, bty = "n", fill = rev(c("brown1", "forestgreen", "dodgerblue4")))


# varpropcum_corr <- lapply(output.vca_corr, function(x) {
#   apply(x, 2, cumsum)
# })
# spectra_summary_corr <- lapply(nirs.listp279_corr, function(x) {
#   data.frame("lambda" = as.integer(colnames(x$NIRS)),
#              "mediane" = apply(x$NIRS, 2, median),
#              "q1" = apply(x$NIRS, 2, function(x){quantile(x, 0.25)}),
#              "q3" = apply(x$NIRS, 2, function(x){quantile(x, 0.75)}))
# })

# #On corrected NIRS
# 
# for (i in 1:length(spectra_summary_corr)){
#   plot(x = spectra_summary_corr[[i]]$lambda, y = spectra_summary_corr[[i]]$mediane, type = "n",
#        ylim = c(0, 100), xlab = "wavelength (nm)",
#        ylab = "Percentage of variances",
#        main = names(varpropcum_corr)[i])
#   polygon(c(spectra_summary_corr[[i]]$lambda, rev(spectra_summary_corr[[i]]$lambda)),
#           c(rep(0, ncol(varpropcum_corr[[i]])), rev(varpropcum_corr[[i]][1, ])),
#           col = "brown1")
#   polygon(c(spectra_summary_corr[[i]]$lambda, rev(spectra_summary_corr[[i]]$lambda)),
#           c(varpropcum_corr[[i]][1, ], rev(varpropcum_corr[[i]][2, ])),
#           col = "forestgreen")
#   polygon(c(spectra_summary_corr[[i]]$lambda, rev(spectra_summary_corr[[i]]$lambda)),
#           c(varpropcum_corr[[i]][2, ], rev(varpropcum_corr[[i]][3, ])),
#           col = "dodgerblue4")
#   par(new = TRUE)
#   plot(spectra_summary_corr[[i]]$lambda, spectra_summary_corr[[i]]$mediane, type = "l",
#        lty = 1, lwd = 1.5, ylim = range(nirs.listp279_corr[[i]]$NIRS),
#        axes = F, xlab = "", ylab = "", main = "", col = "black")
#   polygon(c(spectra_summary_corr[[i]]$lambda, rev(spectra_summary_corr[[i]]$lambda)),
#           c(spectra_summary_corr[[i]]$q1, rev(spectra_summary_corr[[i]]$q3)),
#           col = "#C8C8C888", border = 0)
#   axis(4)
#   mtext("Refelctance", side = 4, line = 2.5, cex = 0.7)
# }
# legend(-100, 0.0019, c("res=bloc/genotype", "bloc", "genotype"),
#        lty = 0, bty = "n", fill = rev(c("brown1", "forestgreen", "dodgerblue4")))
```

Save variance partitions along the spectrum :

```{r save varpropcum}
p2f <- paste0(results.dir,"/spectrum_correction/wood_p279_2020_variance-parts.Rdata")
save(varpropcum, file=p2f)
# save(varpropcum_corr, file=paste0(results.dir,
#                                   "/p279_variance-parts_corr_all_n1-session.Rdata"))
```

## On PCA coordinate axis

```{r pca coord, cache = TRUE}
output.vca.pca <- list()
for (i in 1:length(nirs.list)){
  nirs.list[[i]]$PCs <- pca.ade4[[i]]$li
  output.vca.pca[[i]] <- matrix(NA, nrow = 3, ncol = 10)
  rownames(output.vca.pca[[i]]) <- c("Id", "Id:Bloc", "error")
  colnames(output.vca.pca[[i]]) <- paste0("PC", 1:10)
  for (j in 1:10) {
    tmp <- data.frame(nirs.list[[i]][, c("Id", "Bloc", "Id_bloc")],
                      y = as.vector(nirs.list[[i]]$PCs[, j]))
    vcmod <- anovaVCA(y ~ Id+Bloc, tmp)
    output.vca.pca[[i]][, j] <- vcmod$aov.tab[2:4, "%Total"]
    rm(tmp, vcmod)
  }
  cat(i)
}

# #On corrected NIRS
# 
# output.vca.pca_corr <- list()
# for (i in 1:length(nirs.listp279_corr)){
#   nirs.listp279_corr[[i]]$PCs <- pca.ade4p279_corr[[i]]$li
#   output.vca.pca_corr[[i]] <- matrix(NA, nrow = 3, ncol = 10)
#   rownames(output.vca.pca_corr[[i]]) <- c("Id", "Id:Bloc", "error")
#   colnames(output.vca.pca_corr[[i]]) <- paste0("PC", 1:10)
#   for (j in 1:10) {
#     tmp_corr <- data.frame(nirs.listp279_corr[[i]][, c("Id", "Bloc", "Id_bloc")],
#                       y = as.vector(nirs.listp279_corr[[i]]$PCs[, j]))
#     vcmod_corr <- anovaVCA(y ~ Id+Bloc, tmp_corr)
#     output.vca.pca_corr[[i]][, j] <- vcmod_corr$aov.tab[2:4, "%Total"]
#     rm(tmp_corr, vcmod_corr)
#   }
#   cat(i)
# }
```

```{r percvarcomp pca, fig.height=10, fig.width=10}
par(mfrow = c(3, 2), mar = c(4.5, 4.5, 2, 1))

barplot(t(pve.pca[, 1] * t(output.vca.pca[[1]]) / 100),
          col = c("brown1", "forestgreen", "dodgerblue4"),
          main = colnames(pve.pca)[1],
          ylab="% variance", ylim = c(0, 100))
legend("topright", c("res=bloc/genotype", "bloc", "genotype"),
       lty = 0, bty = "n", fill = rev(c("brown1", "forestgreen", "dodgerblue4")))
for (i in 2:ncol(pve.pca)){
  barplot(t(pve.pca[, i] * t(output.vca.pca[[i]]) / 100),
          col = c("brown1", "forestgreen", "dodgerblue4"),
          main = colnames(pve.pca)[i],
          ylab="% variance", ylim = c(0, 100))
}

# #On corrected NIRS
# 
# barplot(t(pve.pcap279_corr[, 1] * t(output.vca.pca_corr[[1]]) / 100),
#           col = c("brown1", "forestgreen", "dodgerblue4"),
#           main = colnames(pve.pcap279_corr)[1],
#           ylab="% variance", ylim = c(0, 100))
# legend("topright", c("res=bloc/genotype", "bloc", "genotype"),
#        lty = 0, bty = "n", fill = rev(c("brown1", "forestgreen", "dodgerblue4")))
# for (i in 2:ncol(pve.pcap279_corr)){
#   barplot(t(pve.pcap279_corr[, i] * t(output.vca.pca_corr[[i]]) / 100),
#           col = c("brown1", "forestgreen", "dodgerblue4"),
#           main = colnames(pve.pcap279_corr)[i],
#           ylab="% variance", ylim = c(0, 100))
# }
```

# Save Spectra

For phenomic prediction.

Save the mean spectra for each modality (corrected / non corrected and each pre-process).

```{r}
nirs.all <- nirs.list
names(nirs.all)

sp.name <- colnames(nirs.all$raw$NIRS)
X.nirs <- lapply(nirs.all, function(x){
   apply(x$NIRS, 2, function(y) {tapply(y, x$Id, mean)})})
X.nirs$raw[1:5,1:8]

p2f <- paste0(results.dir,"/spectrum_correction/wood_p279_2020_all_spectra.Rdata")
save(X.nirs, file=p2f)
```




# Appendix

```{r info}
stopCluster(cl)
format(Sys.time(), '%d/%m/%Y %H:%M:%S')
t1 <- proc.time(); t1 - t0
print(sessionInfo(), locale=FALSE)
```



